---
title: "Untitled"
format: html
---

Following Jonas game can we make it a shiny app.

### Starting state

Economy is 100
Environment is 10

### Players

3 types of consumers;
- consumes 5
- consumes 3 or 5
- consumes 3

4 types of institutional player
- government decides how much to fund university
- industry funds corporate researcher and decide how they innovate
- university decides how to innovate
- corporate researcher innovates

### Innovations

As university or corporate researcher you take as many cards as you are allocated.

```{r}
library(tidyverse)

cards <- tibble(
  suit = c(
    rep("Clubs", 13),
    rep("Diamonds", 13),
    rep("Hearts", 13),
    rep("Spades", 13)
  ),
  number = rep(c(1:10, "J", "Q", "K"), 4),
  value = rep(c(1:10, rep(10, 3)), 4)
)

```

### Description of a round

```{r}
g_exp = 30

i_exp = 30

uni_inov <- cards %>% sample_n(size = g_exp / 5)

corp_inov <- cards %>% sample_n(size = g_exp / 5)

uni_inov %>%
  count(value) %>%
  filter(n > 1) %>%
  mutate(outcome = value * n) %>%
  pull(outcome)

corp_inov %>%
  count(value) %>%
  filter(n > 1) %>%
  mutate(outcome = value * n) %>%
  pull(outcome)
```


