---
title: "Hisco to Hisclass abbreviation"
format: gfm
---

## Overview

This R project aims to process historical international standard classification of occupations (HISCO) codes to social-economic status (hisclass) levels. The project includes transforming HISCO codes from an Excel spreadsheet into two hisclass levels: 12-level and 7-level classifications, and then mapping these to their respective abbreviated titles.

## Installation

Before running the script, you need to ensure that you have R installed on your system. The following R packages are required:

- `tidyverse`
- `devtools`
- `hisco`
- `readxl`
- `here`

You can install these packages using the following commands in your R console:

```R
install.packages("tidyverse")
install.packages("devtools")
install.packages("readxl")
install.packages("here")
devtools::install_github("junkka/hisco")
```

## Usage

The script can be broken down into several steps:

1. **Setting up the environment:** Loading the necessary libraries.
2. **Creating lookup tables:** Define mappings between HISCO codes and hisclass levels (both 12-level and 7-level).
3. **Reading the data:** Import the HISCO codes from an Excel file.
4. **Processing the data:** Apply the mappings to convert HISCO codes to hisclass levels and join the data with the lookup tables to get the desired outputs.

### Code Breakdown

- **Environment Setup:** The required libraries are loaded into the R environment.
  
- **Lookup Table Creation:** `hisclass_12_tbl` and `hisclass_7_tbl` tibbles are created using predefined mappings to associate numerical hisclass identifiers with their corresponding titles.

- **Data Reading:** HISCO codes are read from an Excel file located in the "data" directory relative to the project root.

- **Data Processing:** HISCO codes are converted to hisclass 12 levels using the `hisco_to_ses()` function. Then the data is joined with `hisclass_table` to populate both hisclass 12 and hisclass 7 levels and their corresponding titles.

## Data Files

The script assumes that there is an Excel file named `hisco_to_hisclass.xlsx` in the `data` directory of your project. The file should contain HISCO codes to be processed.

## Running the Script

To run the script, open the R file in your R environment and execute the code. Ensure that your current working directory is set to the project root and that the `data` folder with the Excel file is correctly placed in the root.


## What will we use?

```{r}
library(tidyverse)

# Data for the first part
hisclass_12_numbers <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)
hisclass_12_labels <- c("Higher managers", "Higher professionals", "Lower managers", 
              "Lower professionals, clerical and sales personnel",
              "Lower clerical and sales personnel", "Foremen", 
              "Medium-skilled workers", "Farmers and fishermen", 
              "Low-skilled workers", "Low-skilled farm workers", 
              "Unskilled workers", "Unskilled farm workers")

# Data for the second part
hisclass_7_numbers <- c(1, 2, 3, 4, 5, 6, 7)
hisclass_7_labels <- c("Elite", "White collar", "Foremen", 
              "Medium-skilled workers", "Farmers and fishermen", 
              "Low-skilled workers", "Unskilled workers")

# Creating the tibbles
hisclass_12_tbl <- tibble(hisclass_12 = hisclass_12_numbers, Title_hisclass_12 = hisclass_12_labels)
hisclass_7_tbl <- tibble(hisclass_7 = hisclass_7_numbers, Title_hisclass_7 = hisclass_7_labels)

hisclass_table <- hisclass_12_tbl %>%
  mutate(cross_walk = case_when(
    hisclass_12 %in% c(1, 2) ~ 1,
    hisclass_12 %in% c(3, 4, 5) ~ 2,
    hisclass_12 %in% c(6) ~ 3,
    hisclass_12 %in% c(7) ~ 4,
    hisclass_12 %in% c(8) ~ 5,
    hisclass_12 %in% c(9, 10) ~ 6,
    hisclass_12 %in% c(11, 12) ~ 7,
  )) %>% 
  inner_join(hisclass_7_tbl, by = c("cross_walk" = "hisclass_7")) %>% 
  rename("hisclass_7" = "cross_walk")



```


## Getting Hisclass from HISCO

```{r}
library(devtools)
install_github("junkka/hisco")
library(hisco)
```

## Read in data from excel

```{r}
library(here)
df <- readxl::read_excel(here("data", "hisco_to_hisclass.xlsx"))

```

## Process

What we want to do is create two more columns - one for the hisclass at 12 level, and one for the the hisclass at 7 level.

To do this, we need to use the `hisco_to_ses(hisco_codes, "hisclass")` function. 

```{r}
df %>% 
  mutate(hisclass_12 = hisco_to_ses(hisco, "hisclass")) %>% 
  inner_join(hisclass_table)
```





