---
title: "Shapefiles"
format: html
execute: 
  echo: true
---

## Purpose

Transform two shapefiles to the same coordinate system and plot them together.

### Getting started

We load the `tidyverse` for data manipulation, the `sf` package fo working with maps and set the ggplot theme to a light one for more attractive plots.

```{r}
library(tidyverse)
library(sf)
theme_set(theme_light())
```

### Read in the shapefiles

We read in our shapefiles with the `st_read()` function, and ask for output of the crs with the `st_crs()` function so that we can see what it is.

```{r, include=FALSE}
library(rnaturalearth)

sweden_map <- ne_countries(country = "sweden", type = "countries", returnclass = "sf", scale = "medium") %>% 
  as_tibble() %>% 
  st_sf() %>% 
  st_transform(3006) %>% 
  st_set_crs(3006)

sweden_map %>% 
  st_write(here::here("data", "sweden_map.shp"), delete_dsn = TRUE)

norway_map <- ne_countries(country = "norway", type = "countries", returnclass = "sf", scale = "medium") %>% 
  as_tibble() %>% 
  st_sf() %>% 
  st_transform(4326) %>% 
  st_set_crs(4326)

norway_map %>% 
  st_write(here::here("data", "norway_map.shp"), delete_dsn = TRUE)
```

For Sweden we can see that it is **SWEREF99 TM** a projection favoured by the statistics body in Sweden.

```{r}
sweden_map <- st_read(here::here("data", "sweden_map.shp"))

sweden_map %>% 
  st_crs()
```

For Norway it is **WGS 84**, a very common projection with full name **World Geodetic System 1984**.

```{r}
norway_map <- st_read(here::here("data", "norway_map.shp"))

norway_map %>% 
  st_crs()
```

Let's have a quick look at them side by side.

```{r}
#| layout-ncol: 2
#| fig-cap: 
#|   - "Sweden"
#|   - "Norway"

ggplot()+ 
  geom_sf(data = sweden_map) +
  labs(title = "Sweden",
       subtitle = "CRS = SWEREF99")

ggplot()+ 
  geom_sf(data = norway_map) +
  labs(title = "Norway",
       subtitle = "CRS = WGS84")
```


### Set the crs as the same

We'll use EPSG 4326 or WGS84. We can use the `st_transform()` function, specifying the crs that we want for our new map object, which we name `sweden_map_4326` to avoid confusion.

This is what the CRS now looks like for Sweden - the same as the one for Norway.

```{r}
sweden_map_4326 <- st_transform(sweden_map, crs = 4326)

sweden_map_4326 %>% 
  st_crs()
```

### Let's plot them together

We can plot the together and choose some nice colours from the national flags for the fill aesthetic.

```{r}
ggplot() +
  geom_sf(data = sweden_map_4326, fill = "#03558A") +
  geom_sf(data = norway_map, fill = "#B40C2E") +
  labs(title = "")
```

### Write out the file for Sweden individually

We can save the shapefile for Sweden with the `st_write()` command if we want to keep the maps separate.

```{r}
sweden_map_4326 %>% 
  st_write(here::here("data", "sweden_map_4326.shp"), delete_dsn = TRUE)
```

### Join the maps together

Alternatively we can join together the maps as one object and then save one shapefile with the maps combined. 

To do this, we first bind the two objects together using `bind_rows()` command and write it out with the same `st_write()` command.

```{r}
combined_map <- bind_rows(norway_map, sweden_map_4326)

combined_map %>% st_write(here::here("data", "shapefile_combined.shp"), delete_dsn = TRUE)
```

If we want to get rid of the borders between our objects, we can use the `st_union()` command:

```{r}
union_map <- st_union(norway_map, sweden_map_4326)
```

We can compare the outputs here:

```{r}
#| layout-ncol: 2
#| fig-cap: 
#|   - "Output from `bind_rows`"
#|   - "Output from `st_union`"

ggplot() +
  geom_sf(data = combined_map, aes(fill = sovereignt)) +
  scale_fill_manual(values = c("#B40C2E", "#03558A")) +
  labs(fill = "Country")

ggplot() +
  geom_sf(data = union_map, aes(fill = sovereignt)) +
  scale_fill_manual(values = c("darkgreen")) +
  labs(fill = "Country")
```

### Conclusion

Here we learned how to make it possible to plot maps with two different projections together by making use of the `st_transform()` command in R. We learned how to save the two maps as separate features in a shapefile, as well as how to combine them togther.

